### 2016-09-01 ms/ms-bbl.tex
Added coverletter to repo

\begin{thebibliography}{10}

\bibitem{Croft2015}
R.~Croft, T.~{Di Matteo}, and N.~Khandai, ``{Petascale Cosmology: Simulations
  of Structure Formation},'' {\em Comput. Sci. Eng.}, vol.~17, pp.~40--46, mar
  2015.

\bibitem{Juric2015a}
M.~Juri{\'{c}}, J.~Kantor, K.-T. Lim, R.~H. Lupton, G.~Dubois-Felsmann,
  T.~Jenness, T.~S. Axelrod, J.~Aleksi{\'{c}}, R.~A. Allsman, Y.~AlSayyad,
  J.~Alt, R.~Armstrong, J.~Basney, A.~C. Becker, J.~Becla, S.~J. Bickerton,
  R.~Biswas, J.~Bosch, D.~Boutigny, M.~{Carrasco Kind}, D.~R. Ciardi, A.~J.
  Connolly, S.~F. Daniel, G.~E. Daues, F.~Economou, H.-F. Chiang, A.~Fausti,
  M.~Fisher-Levine, D.~M. Freemon, P.~Gee, P.~Gris, F.~Hernandez, J.~Hoblitt,
  {\v{Z}}.~Ivezi{\'{c}}, F.~Jammes, D.~Jevremovi{\'{c}}, R.~L. Jones, J.~{Bryce
  Kalmbach}, V.~P. Kasliwal, K.~S. Krughoff, D.~Lang, J.~Lurie, N.~B. Lust,
  F.~Mullally, L.~A. MacArthur, P.~Melchior, J.~Moeyens, D.~L. Nidever,
  R.~Owen, J.~K. Parejko, J.~M. Peterson, D.~Petravick, S.~R. Pietrowicz, P.~A.
  Price, D.~J. Reiss, R.~A. Shaw, J.~Sick, C.~T. Slater, M.~A. Strauss, I.~S.
  Sullivan, J.~D. Swinbank, S.~{Van Dyk}, V.~Vuj{\v{c}}i{\'{c}}, A.~Withers,
  P.~Yoachim, and {LSST Project}, ``{The LSST Data Management System},'' {\em
  eprint arXiv:1512.07914}, 2015.

\bibitem{Morton1996}
G.~M. Morton, ``{A computer oriented geodetic data base and a new technique in
  file sequencing},'' tech. rep., IBM Ltd., Ottawa, Ontario, 1996.

\bibitem{Hilbert1970}
D.~Hilbert, {\em {Gesammelte Abhandlungen: Band III: Analysis - Grundlagen der
  Mathematik Physik - Verschiedenes Lebensgeschichte}}, ch.~{\{}{\"{U}}{\}}ber
  die, pp.~1--2.
\newblock Berlin, Heidelberg: Springer Berlin Heidelberg, 1970.

\bibitem{Springel2005b}
V.~Springel, ``{The cosmological simulation code gadget-2},'' {\em Mon. Not. R.
  Astron. Soc.}, vol.~364, pp.~1105--1134, dec 2005.

\bibitem{Teyssier2001}
R.~Teyssier, ``{Cosmological Hydrodynamics with Adaptive Mesh Refinement: a new
  high resolution code called RAMSES},'' {\em Astron. Astrophys. v.385,
  p.337-364}, vol.~385, pp.~337--364, nov 2001.

\bibitem{Hjaltason2002}
G.~R. Hjaltason and H.~Samet, ``{Speeding up construction of PMR quadtree-based
  spatial indexes},'' {\em VLDB J. Int. J. Very Large Data Bases}, vol.~11,
  pp.~109--137, oct 2002.

\bibitem{Connor2010}
M.~Connor and P.~Kumar, ``{Fast construction of k-nearest neighbor graphs for
  point clouds.},'' {\em IEEE Trans. Vis. Comput. Graph.}, vol.~16,
  pp.~599--608, jan 2010.

\bibitem{Orenstein1984}
J.~A. Orenstein and T.~H. Merrett, ``{A class of data structures for
  associative searching},'' in {\em Proc. 3rd ACM SIGACT-SIGMOD Symp. Princ.
  database Syst. - Pod. '84}, (New York, New York, USA), p.~181, ACM Press, apr
  1984.

\bibitem{Wu1998}
M.-C. Wu and A.~Buchmann, ``{Encoded bitmap indexing for data warehouses},'' in
  {\em Proc. 14th Int. Conf. Data Eng.}, pp.~220--230, IEEE Comput. Soc, 1998.

\bibitem{Chan1998}
C.-Y. Chan and Y.~E. Ioannidis, ``{Bitmap index design and evaluation},'' {\em
  ACM SIGMOD Rec.}, vol.~27, pp.~355--366, jun 1998.

\bibitem{Chan1999}
C.-Y. Chan and Y.~E. Ioannidis, ``{An efficient bitmap encoding scheme for
  selection queries},'' {\em ACM SIGMOD Rec.}, vol.~28, pp.~215--226, jun 1999.

\bibitem{Malensek2014}
M.~Malensek, S.~Pallickara, and S.~Pallickara, ``{Evaluating Geospatial
  Geometry and Proximity Queries Using Distributed Hash Tables},'' {\em Comput.
  Sci. Eng.}, vol.~16, pp.~53--61, jul 2014.

\bibitem{Sinha2006}
R.~Sinha, S.~Mitra, and M.~Winslett, ``{Bitmap indexes for large scientific
  data sets: a case study},'' in {\em Proc. 20th IEEE Int. Parallel Distrib.
  Process. Symp.}, p.~10 pp., IEEE, 2006.

\bibitem{Sinha2007}
R.~R. Sinha and M.~Winslett, ``{Multi-resolution bitmap indexes for scientific
  data},'' {\em ACM Trans. Database Syst.}, vol.~32, pp.~16--es, aug 2007.

\bibitem{Stockinger2000}
K.~Stockinger, D.~Duellmann, W.~Hoschek, and E.~Schikuta, ``{Improving the
  Performance of High-Energy Physics Analysis through Bitmap Indices},'' in
  {\em Database Expert Syst. Appl. 11th Int. Conf. DEXA 2000 London, UK, Sept.
  4--8, 2000 Proc.} (M.~Ibrahim, J.~K{\"{u}}ng, and N.~Revell, eds.), vol.~1873
  of {\em Lecture Notes in Computer Science}, ch.~Improving, pp.~835--845,
  Berlin, Heidelberg: Springer Berlin Heidelberg, jun 2000.

\bibitem{Wu2003}
K.~Wu, W.~Koegler, J.~Chen, and A.~Shoshani, ``{Using bitmap index for
  interactive exploration of large datasets},'' in {\em 15th Int. Conf. Sci.
  Stat. Database Manag. 2003.}, pp.~65--74, IEEE Comput. Soc, 2003.

\bibitem{Yu1998}
P.~Yu, ``{Range-based bitmap indexing for high cardinality attributes with
  skew},'' in {\em Proceedings. Twenty-Second Annu. Int. Comput. Softw. Appl.
  Conf. (Compsac '98) (Cat. No.98CB 36241)}, pp.~61--66, IEEE Comput. Soc,
  1998.

\bibitem{Shoshani1999}
A.~Shoshani, L.~Bernardo, H.~Nordberg, D.~Rotem, and A.~Sim,
  ``{Multidimensional indexing and query coordination for tertiary storage
  management},'' in {\em Proceedings. Elev. Int. Conf. Sci. Stat. Database
  Manag.}, pp.~214--225, IEEE Comput. Soc, 1999.

\bibitem{Stockinger2004}
K.~Stockinger, K.~Wu, and A.~Shoshani, ``{Evaluation Strategies for Bitmap
  Indices with Binning},'' in {\em Database Expert Syst. Appl. 15th Int. Conf.
  DEXA 2004, Zaragoza, Spain, August 30-September 3, 2004. Proc.} (F.~Galindo,
  M.~Takizawa, and R.~Traunm{\"{u}}ller, eds.), (Berlin, Heidelberg),
  pp.~120--129, Springer Berlin Heidelberg, 2004.

\bibitem{Wu2001}
K.~Wu, E.~Otoo, and A.~Shoshani, ``{Compressed bitmap indices for efficient
  query processing},'' {\em Lawrence Berkeley Natl. Lab.}, sep 2001.

\bibitem{Lemire2010}
D.~Lemire, O.~Kaser, and K.~Aouiche, ``{Sorting improves word-aligned bitmap
  indexes},'' {\em Data Knowl. Eng.}, vol.~69, pp.~3--28, jan 2010.

\bibitem{Kaser2016}
O.~Kaser and D.~Lemire, ``{Compressed bitmap indexes: beyond unions and
  intersections},'' {\em Softw. Pract. Exp.}, vol.~46, pp.~167--198, feb 2016.

\bibitem{Monaghan1992}
J.~J. Monaghan, ``{Smoothed Particle Hydrodynamics},'' {\em Annu. Rev. Astron.
  Astrophys.}, vol.~30, pp.~543--574, sep 1992.

\bibitem{Springel2001}
V.~Springel, N.~Yoshida, and S.~D.~M. White, ``{GADGET: a code for
  collisionless and gasdynamical cosmological simulations},'' {\em New
  Astron.}, vol.~6, pp.~79--117, apr 2001.

\bibitem{Hopkins2015}
P.~F. Hopkins, ``{A new class of accurate, mesh-free hydrodynamic simulation
  methods},'' {\em Mon. Not. R. Astron. Soc.}, vol.~450, pp.~53--110, apr 2015.

\bibitem{Turk20d11a}
M.~J. Turk, B.~D. Smith, J.~S. Oishi, S.~Skory, S.~W. Skillman, T.~Abel, and
  M.~L. Norman, ``{yt: A MULTI-CODE ANALYSIS TOOLKIT FOR ASTROPHYSICAL
  SIMULATION DATA},'' {\em Astrophys. J. Suppl. Ser.}, vol.~192, p.~9, jan
  2011.

\bibitem{Behnel2011}
S.~Behnel, R.~Bradshaw, C.~Citro, L.~Dalcin, D.~S. Seljebotn, and K.~Smith,
  ``{Cython: The Best of Both Worlds},'' {\em Comput. Sci. Eng.}, vol.~13,
  pp.~31--39, mar 2011.

\end{thebibliography}